---
title: "EDS 213 Section Project: Query and Plot Outputs"
author: "Tom Gibbens-Matsuyama"
format: html
editor: visual
---

### Section Project

In this Quarto document, I will be demonstrating the query that was used SQL by reproducing it with Dbplyr code. The data used in this project is crop data consisting of countries crop yields and other related variables.

### Load libraries

```{r}
library(DBI) 
library(tidyverse)
library(duckdb)
library(dbplyr)
```

### Load data

```{r}

# Connect to database
conn <- DBI::dbConnect(duckdb::duckdb(), dbdir='database.db')
DBI::dbListTables(conn)

# Read in database data
crop_yield <- tbl(conn, "crop_yield")

crop_calendar <- tbl(conn, "crop_calendar")
```

### Let's go through the workflow I created in SQL

### My Questions:

- My question: What are the countries for the top 5 yields for each crop? 
- My question: What is the minimum, maximum, and mean average_temp and average_precip for the top 5 yields of each crop? 

In my query, I filtered down my data a little more before joining together. I filtered the crop_yield column to only having yield outputs for the `elements` column

```{r}
# Filter so we only have a yield observation
crop_yield <- crop_yield %>% 
  filter(element == "Yield")
```

### Join the two tables

```{r}
# Join our two tables together
crop_join <- inner_join(crop_calendar, crop_yield, by = c("country", "crop"))
```

### Question 1:
For my first question I want to find what countries make up the top 5 yields for each crop. This means I should have 25 observations, as I want to calculate the top 5 yields for 5 crops. I will use the same workflow as my SQL query. 

```{r}
# Get rid of duplicate country observations, if there are any
unique_yields <- crop_join %>% 
  group_by(crop, value) %>% 
  summarise(country = max(country)) %>% 
  ungroup()

# Create a ranking of the countries by yield value
ranked_yields <- unique_yields %>% 
  group_by(crop) %>% 
  mutate(rank = min_rank(desc(value))) %>% 
  ungroup()

# Filter to only top 5 ranking per crop 
top5_yields <- ranked_yields %>% 
  filter(rank <=5) %>% 
  select(crop, value, country)
```

### Question 2: 
My question: What is the minimum, maximum, and mean temp_average and precip_average for the top 5 yields of each crop? Again, I will use the same workflow and naming conventions as my SQL query. 

```{r}
# Creating a new top 5 yields column
top5_yields_2 <- ranked_yields %>% 
  filter(rank <=5)

# Join back with crop data to get avg temp and precip data
top5_climate <- top5_yields_2 %>% 
  inner_join(crop_join, by = c("crop", "value", "country"))

# I can do an inner join to get the other variables back
climate_summary_stats <- top5_climate %>% 
  group_by(crop) %>% 
  summarise(
    min_temp = round(min(temp_average, na.rm = TRUE), 2),
    max_temp = round(max(temp_average, na.rm = TRUE), 2),
    mean_temp = round(mean(temp_average, na.rm = TRUE), 2),
    min_precip = round(min(precip_average, na.rm = TRUE), 2),
    max_precip = round(max(precip_average, na.rm = TRUE), 2),
    mean_precip = round(mean(precip_average, na.rm = TRUE), 2),
    .groups = "drop"
  ) 

# For tempareature data: need to pivot longer for plotting
climate_pivot_temp <- climate_summary_stats %>% 
  select(crop, min_temp, max_temp, mean_temp) %>% 
  pivot_longer(cols = -crop,
               names_to = "summary_stat",
               values_to = "value")


# Plotting the temperature data
ggplot(climate_pivot_temp, aes(x = crop, y = value, fill = summary_stat)) +
  geom_col(position = "dodge") +
  labs(x = element_blank(),
       y = "Temperature (Degrees Celsius)",
       title = "Temperature Summary Stats for the Top 5 Crop Yields") +
  theme_minimal()


# For precipitation data: need to pivot longer for the plot
climate_pivot_precip <- climate_summary_stats %>% 
  select(crop, min_precip, max_precip, mean_precip) %>% 
  pivot_longer(cols = -crop,
               names_to = "summary_stat",
               values_to = "value")


# Plotting the precipitaiton data
ggplot(climate_pivot_precip, aes(x = crop, y = value, fill = summary_stat)) +
  geom_col(position = "dodge") +
  labs(x = element_blank(),
       y = "Precipitation (mm per Month)",
       title = "Precipitation Summary Stats for the Top 5 Crop Yields") +
  theme_minimal()
```

